Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-19 paper/paper.tex
illustrative rl.png

\author{Elvis Dohmatob\textsuperscript{1,2}, Guillaume Dumas\textsuperscript{3,4,5,6}, Danilo Bzdok\textsuperscript{1,2}
}

\maketitle
\small{
    \textbf{1} INRIA, Parietal team, Saclay, France\\
    \textbf{2} CEA, Neurospin, Gif-sur-Yvette, France\\
    \textbf{3} Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
    \textbf{4} CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
    \textbf{5} University Paris Diderot, Sorbonne Paris Cité, Paris, France\\
    \textbf{6} Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
}

\begin{abstract}
The default mode network (DMN) is believed to subserve the human
baseline mental activity.
%
Its highest energy consumption in the brain and
its intimate coupling with conscious awareness are both pointing towards an important functional roles.
%
Many research streams support its evolutionarily adaptive role as
envisioning mental scenarios to simulate the future.

The present concept paper proposes to explain \textit{how} the DMN may actually implement such continuous assessment and prediction of the environment to guide action choices.

DMN function is recast in mathematical terms of
control theory and reinforcement learning based on Markov decision processes.
%
We argue that our formal account of DMN function naturally accommodates as special cases
previously proposed cognitive accounts such as \textit{predictive coding}.
%
A process model for the neural implemention
of continuous future predictions potentially realized in the DMN
could offer insight into the statistical mechanisms underlying human behavior and provide a parsimonious explanation of experimental findings in animals and humans.
\end{abstract}

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain
neuronal computation
in the absence of behavioral focus
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is surprisingly weakly modulated
with task demand \citep{raichle2001pnas}. At the opposite, during sleep, the
decoupling of brain structures discarded the idea of the DMN being only a passive network resonance and rather supported its important role in the sustenance of conscious awareness \citep{horovitz2009decoupling}.
This dark matter of brain physiology begs the question
of underlying biological purpose
\citep{raichle2006brain}.
What has early been described as the "stream of consciousness"
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the so-called "default mode network" (DMN)
\citep{shul1997, raichle2001pnas}.
This set of some of the highest regions in
the human association cortex \citep{mesulam1998sensation, margulies2016situating}
consistently
increase in neural activity during
unfocused everyday mind wandering \citep{raichle2001pnas}.
The DMN may subserve important functions in
higher-order control of human behavior
following the notion of "a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex" \citep{carhart2010default}.
In the beginning of the 21st century,
brain imaging may have been the first technique
to allow for the discovery of a unique brain network that subserves
baseline mental activities
\citep{raichle2001pnas, bzdok2015resting}.

\subsection{Default mode network: higher-order control of the organism}
The human DMN is responsible for a large part of the brain's baseline activity,
which typically decreases during psychological experiments.
The standard mode of
neural information maintenance and manipulation has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
In line with this, the DMN contains hotspots
of highest energy consumption in the entire central nervous system
\citep{raichle2001pnas},
which persists to a substantial degree during
sleep and under anesthesia \citep{randy2008}.
Today, many authors believe that the DMN implements
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008}.
This brain network
might have emerged to continuously predict environmental events using
mental imagery as an evolutionary advantage.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode areas \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into the function of the DMN is
particularly challenging because
this network appears to
simultaneously control perception-action cycles and
maintain baseline contemplations
across time, space, and content domains.

The present work takes the perspective of
a human agent faced with the choice of the next action
and being guided by rich
reenactment of really-happened, hypothetically conceived, and
estimated futures to optmize behavioral performance.
A particularly attractive computational framework
to describe, quantify, and predict autonomously acting systems like the brain
is proposed to be the combination of control theory and
reinforcement learning (RL).
It is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation}.
Conversely,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
These ``offline'' processes may however contribute to optimized control of the organism. \textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment? - danilo: hmm, let's keeping thinking about this)}
% We formalize
% a policy matrix to capture the space of possible actions that can be performed
% on the environment given the current state. A value matrix
% maps expected rewards to environmental objects and events.
% Switching between states reduces to a sequential processing model.
% In a model-based reinforcement learning (RL) approach,
% we try to understand the environment always better
% starting, at birth, with an environment that is completely unknown.
Informed by outcomes of performed action,
the DMN dynamics are constantly adapted in feedback loops
shaped by prediction error.
A DMN framework incorporating RL can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term reward outcomes.

\subsection{Towards a formal account of DMN function}
We argue that DMN implication in a
diversity of humans' most advanced cognitive processes
can be parsimoniously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing action outcome across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.
In this context,
\textit{control} refers to the influence that an agent exerts when interacting
with the environment to encourage preferred states.
Even (necessarily imperfect) memory recall,
random day-time mind-wandering, and dreams during sleep
may provide meaningful building blocks to iteratively improve
the predictive DMN machinery to optimize the behavior of the organism.
%
It has indeed been proposed earlier that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
This idea has been invigorated by
empirical evidence from
neuroscientific experiments \citep{kording2004bayesian, fiser2004small}.
The present paper proposes a
process model that satisfies this contention.

\section{Known neurobiological properties of the default mode network}
We begin by a neurobiological deconstruction of the DMN
based on experimental findings in the neuroscience literature.
This walkthrough across network nodes will
outline the individual functional profiles,
paving the way for their functional interpretation
within our formal account (next section).

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{Neurobiological overview of the DMN ... \textbf{Please add a figure caption}}
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are probably responsible for the highest turn-over
of glucose energy consumption \citep{raichle2001pnas}.
These metabolic characteristics go hand-in-hand with
functional neuroimaging analyses \citep{andrews2010} in
suggesting that the PMC and mPFC
potentially represent the functional backbone of the DMN.

The PMC myelinates relatively late during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
Normal and disturbed metabolic fluctuations in the
human PMC have been repeatedly related to
phenomena of changed conscious awareness \citep{cavanna2006precuneus}.
%
Indeed,
the PMC has long been speculated to reflect constant computation of
environmental statistics and its internal representation
as an inner "mind's eye" \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to combine various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.

The putative role of the PMC in continuous integration of
relevant environmental information
and ensuing top-level control of action on the environment is supported
by a number of neuroscience experiments.
Electrophysiological recordings in animals implicated the PMC in
strategic decision making \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
as well as approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would persue an exploratory or exploitative
behavioral strategy during food foraging \citep{pearson2009neurons}.
Further, single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.
This DMN node encoded the
preference of or aversion to options with uncertain reward outcomes
and its spiking activity was more associated with
subjectively perceived relevance of a chosen object
than by its factual value,
based on an internal currency of value \citep{mccoy2005risk}.
In fact, direct stimulation of PMC neurons
promoted exploratory actions,
which would otherwise be shunned \citep{hayden2008stim}.
Graded changes in firing rates of PMC neurons
indicated changing choices in upcoming trials and their neural patterns were
distinct from spike firings that indicated choosing either option.
Similarly in humans,
the DMN has been shown to gather and integrate information
over different parts of auditory narratives in an fMRI study
\citep{simony2016dynamic}.

The retrosplenial portion of the PMC can subserve
representation of action possibilities
and evaluation of reward contingencies by integration with
information from memory and different perspective frames.
Regarding memory recall, retrosplenial lesions have been
consistently associated with anterograde and retrograde memory impairments
of various kinds of sensory information
in rabbits and humans
\citep{vann2009does}.
Regarding perspective frames, the retrosplenial subregion of the PMC has been
proposed to mediate between the organism's egocentric
(i.e., focused on sensory input) and
allocentric (i.e., focused on world knowledge) viewpoints
in animals and humans
\citep{epstein2008parahippocampal, burgess2008spatial, valiquette2007different}.

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible decisions and integrating that information
with memory, perspective frames, and
reward outcomes into higher-level strategies.
Perceived value, that differs across individuals, may update
statistical assessment of the environment
to predict delayed reward opportunities in the future.
In doing so, the PMC may continuously adapt to changes
in both the external environment and its internal representation
to enable strategic behavior in volatile environments.

\subsection{The prefrontal cortex: stimulus-value association and action consideration}
Analogous to the PMC,
the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
supra-sensory processes
across time, space, and content domains to
exert top-down control.
In comparison to the PMC, however,
dmPFC function may be closer to a
``mental sketchpad'' \citep{goldman1996prefrontal}, as it
potentially subserves the de-novo generation and manipulation
of meaning representations instructed by stored semantics and memories.

Generally,
patients with neurological lesions in the prefrontal cortex
are known to struggle with
adaption to novel situations and stimuli
\citep{stuss1986frontal}.
The dmPFC may subserve inference, representation, and assessment
of one's own and other individuals' action considerations.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
errors in these predictions.
Neural activity in the dmPFC indeed explained the proficiency decline
of inferring other peoples' thoughts in aging humans \citep{moran2012social}.
Some dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
In fact, the topographically neighboring dorsal anterior cingulate cortex
has also been linked to computing values and efforts of
persisting a behavioral plan versus switching the
environmental context in several lesion studies \citep{kolling2016value}.
%
Such highly abstract neural computations necessarily rely on the
generation of probabilistic internal information drawing from
episodic memory recall, generative construction processes,
and explicit knowledge of the external world.
%
In a computational neuroimaging experiment,
dorsomedial PFC activity preferentially modeled the consequences of
action choices that were not actually executed,
whereas ventromedial PFC (vmPFC) activity
modeled especially value outcomes of performed actions in the environment
\citep{nicolle2012agent}.

Comparing to the dmPFC,
the vmPFC probably subserves
subjective value evaluation and
risk estimation of relevant environmental stimuli.
This DMN node is more closely associated with
orchestrating adaptive behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on sophisticated value representations
\citep{doherty2015structure}.
Quantitative lesion findings across 344 human individuals confirmed
a substantial impairment in value-based decision making
\citep{glascher2012lesion}.
Indeed,
the vmPFC is preferentially connected with reward-related and limbic areas.
The vmPFC is well known to have monosynaptical connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Congruently, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
in humans
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals from dopaminergic pathways
\citep{schultz1998predictive}
that not only channel action towards basic survival needs but also
subserve RL in humans more broadly \citep{doherty2015structure}.
This is consistent with diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative} that
quantified the NAc to
be more connected to the vmPFC than dmPFC in both species.
Two functional connectivity analyses in humans strongly connected
the vmPFC with the NAc, hippocampus (HC),
and PMC.
%
In line with these connectivity findings in animals and humans,
the vmPFC is often proposed to be involved in (external) emotional
reactions and own (visceral) arousal \citep{damasio1996somatic}.
Real or imagined bodily states could be mapped in the vmPFC
as a bioregulatory disposition governing cognition
and decision making.
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
valuation
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in the vmPFC.
Such a ``cognitive map'' of the action space was argued to encode
the current task state even when states are unobservable from sensory input,
which was shown to be critical for behavior \citep{Schuck2016}.

@@

While the HC
is traditionally believed to allow remembering the past,
there is now increasing evidence for a role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017}.
Its recursive anatomical architecture
may be specifically designed to allow reconstructing
entire episodes of experience from memory fragments.
Indeed,
hippocampal damage is
not only associated with an impairment in reexperiencing the past (i.e., amnesia),
but also imagining one's own future and
fictitious experiences more broadly \citep{hassabis2007patients}.
Mental scenes created by HC-lesioned patients exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal HC revealed
some constantly active neuronal ensembles whose firing coincided with
specific locations in space during environmental navigation.
London taxi drivers, individuals with high performance in spatial navigation,
were shown to exhibit increased gray-matter volume in the
HC \citep{maguire2000navigation}.
Indeed, when an animal is choosing between alternative
paths, the corresponding neuronal populations in the HC
spike one after another  \citep{johnson2007neural}.
These neuronal patterns appear to also indicate upcoming behavior
\citep{pfeiffer2013hippocampal}.

HC function probably extend beyond simple forms of
encoding and reconstruction of memory and spatial information.
Based on large-scale recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, which were elicited in experimental task conditions,
have been shown to be reenacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, spike sequences measured in hippocampal place cells of rats
featured re-occurrence directly after experimental trials
as well as directly before upcoming experimental trials \citep{diba2007forward}.
Such hippocampal ensembles burt during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially also in the nodes of the DMN.
These HC-subserved mechanisms
probably contribute to advanced cognitive processes that require
recollecting autobiographical memory episodes, as well as
reexperiencing or newly generating mental scenarios
\citep{hassabis2007patients}.

>>>>

\author{Elvis Dohmatob, Guillaume Dumas, Danilo Bzdok\\
  \textbf{Please add your affiliation}
  %% INRIA, Parietal team, Saclay, France\\
  %% CEA, Neurospin, Gif-sur-Yvette, France\\
  %% firstname.lastname@inria.fr
  %% GD1 Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France
  %% GD2 CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France
  %% GD3 University Paris Diderot, Sorbonne Paris Cité, Paris, France
  %% GD4 Centre de Bioinformatique, Biostatistique et Biologie Intégrative (C3BI, USR 3756 Institut Pasteur et CNRS), Paris, France
}

\maketitle

\begin{abstract}
The default mode network (DMN) is believed to subserve the human
baseline mental activity.
%
Many research streams converge to its likely role in evolutionarily adaptive
envisioning of mental scenarios to simulate the future.
Such functional roles become more likely due to
this brain network's highest energy consumption in the brain and
its intimate coupling with conscious awareness.
%
The present concept paper proposes a \textit{process model}
in trying to explain \textit{how} the DMN may actually implement continuous
assessment and prediction of the environment to guide action choices.
DMN function is recast in the mathematical terms of
control theory and reinforcement learning based on Markov decision processes.
We argue that our formal account of DMN function
naturally accommodates
previously proposed cognitive accounts as special cases and
parsimoniously explains
experimental findings in animals and humans.
%
A process models for the neural implemention
of continuous future predictions probably realized in the DMN
could offer insight into the statistical mechanisms underlying human behavior.
\end{abstract}

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain
neuronal computation
in the absence of behavioral focus
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is subject to surprisingly little modulations
when mentally processing environmental challenges
\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep? Danilo: I'd love to cite that
but I am not aware of such studies)}
This dark matter of brain physiology begs the question
of underlying biological purpose
\citep{raichle2006brain}.
What has early been described as the "stream of consciousness"
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the so-called "default mode network" (DMN)
\citep{shul1997, raichle2001pnas}.
This set of some of the highest regions in
the human association cortex \citep{mesulam1998sensation, margulies2016situating}
consistently
increased in neural activity during
unfocused everyday mind wandering \citep{raichle2001pnas}.
The DMN may exert higher-order control on human behavior
following the notion of "a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex" \citep{carhart2010default}.
In the beginning of the 21st century,
brain imaging may have been the first technique
to allow for the discovery of a unique brain network that subserves
baseline mental activities
\citep{raichle2001pnas, bzdok2015resting}.

\subsection{Default mode network: higher-order control of the organism}
The human DMN accounts for a large part of the brain's baseline activity
and typically decreases in activity during psychological experiments.
The standard mode of
neural information maintenance and manipulation has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
This is all the more likely because the DMN contains hotspots
of highest energy consumption in the entire central nervous system
\citep{raichle2001pnas}.
DMN activity also persists to a substantial degree during
sleep and under anesthesia \citep{randy2008}.
Today, many authors believe that the DMN implements
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008}.
This brain network
might have emerged to continuously predict environmental events using
mental imagery as an evolutionary advantage.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode areas \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into the overarching DMN function is challenging because
this network appears to
simultaneously control perception-action cycles and
maintain baseline contemplations
across time, space, and content domains.

In this work,
we consider an agent faced with the choice of the next action
and being guided by rich
reenactment of really-happened, hypothetically conceived, and
predicted future events to perfect ongoing behavioral performance.
A particularly attractive computational framework
to describe, quantify, and predict autonomously acting systems like the brain
is proposed to be the combination of control theory and
reinforcement learning (RL).
It is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation}.
Conversely,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
These ``offline'' processes may however contribute to optimized control of the organism. \textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment? - danilo: hmm, let's keeping thinking about this)}
% We formalize
% a policy matrix to capture the space of possible actions that can be performed
% on the environment given the current state. A value matrix
% maps expected rewards to environmental objects and events.
% Switching between states reduces to a sequential processing model.
% In a model-based reinforcement learning (RL) approach,
% we try to understand the environment always better
% starting, at birth, with an environment that is completely unknown.
Informed by outcomes of performed action,
the DMN dynamics are constantly adapted in feedback loops
shaped by prediction error.
A DMN framework incorporating RL can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term reward schedules.

\subsection{Towards a formal account of DMN function}
We argue that the implication of the DMN in a
diversity of humans' most advanced cognitive processes
can be parsimoniously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing reward outcome across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
\citep{tenenbaum2011grow}.
In this context,
\textit{control} refers to the influences that an agent exerts when interacting
with the environment to encourage preferred states.
Even (necessarily imperfect) memories, random day-time mind-wandering, and dreams during sleep
may provide meaningful building blocks to iteratively improve
the predictive DMN machinery to optimize the behavior of the organism.
%
In short, it has been proposed earlier that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
This idea has been invigorated by
empirical evidence from
neuroscientific experiments \citep{kording2004bayesian, fiser2004small}.
The present paper proposes a
process model that satisfies this contention.

\section{Known neurobiological properties of the default mode network}
We begin by a neurobiological deconstruction of the DMN
based on experimental findings in the neuroscience literature.
This walkthrough across network nodes will
outline their individual functional profiles,
paving the way for their functional interpretation within our computational account
described in the next section.

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{Neurobiological overview of the DMN ... \textbf{Please add a figure caption}}
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are responsible for the highest turn-over
of glucose energy consumption \citep{raichle2001pnas}.
These metabolic characteristics go hand-in-hand with
brain-imaging analyses \citep{andrews2010} in
suggesting that the PMC and mPFC
potentially represent the functional backbone of the DMN.

The PMC myelinates relatively late during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
Normal and disturbed metabolic fluctuations in the
human PMC have been repeatedly related to
phenomena of changed conscious awareness \citep{cavanna2006precuneus},
including anesthesia, sleep, and coma.
%
Indeed,
the PMC has long been speculated to reflect constant computation of
environmental statistics and its internal representation
as an inner "mind's eye" \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
the inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.

The putative role of the PMC in continuous integration of
relevant environmental information
and ensuing top-level control of action on the environment is supported
by a number of neuroscience experiments.
Electrophysiological recordings in animals implicated the PMC in
strategic decision making \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
and outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
while its retrosplenial portion was
repeatedly implicated in approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would persue an exploratory or exploitative
behavioral strategy during food foraging \citep{pearson2009neurons}.
Further, single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.
This DMN node encoded the
preference of or aversion to options with uncertain reward outcomes
and its spiking activity was more associated with
subjectively perceived relevance of a chosen object
than by its factual value,
based on an internal currency of value \citep{mccoy2005risk}.
In fact, direct stimulation of PMC neurons
promoted exploratory action towards actions
that were previously shunned \citep{hayden2008stim}.
Graded changes in firing rates of PMC neurons
indicated changing choices in upcoming trials and their neural patterns were
distinct from spike firings that indicated choosing either option.
Similarly in humans,
the DMN has been shown to gather and integrate information
over different paragraphs in an fMRI study on auditory narratives
\citep{simony2016dynamic}.

The retrosplenial portion of the PMC can subserve action possibilities
and evaluation of reward contingencies by integrating these with
information from memory and different perspective frames.
Regarding memory retrieval, retrosplenial lesions have been
consistently associated with anterograde and retrograde memory impairments
of various kinds of sensory information
in rabbits and humans
\citep{vann2009does}.
Regarding perspective frames, this PMC subregion has been
proposed to mediate between the organism's egocentric
(i.e., focused on sensory input) and
allocentric (i.e., focused on world knowledge) viewpoints
in animals and humans
\citep{epstein2008parahippocampal, burgess2008spatial, valiquette2007different}.

Consequently, the PMC may contribute to overall DMN function
by monitoring the subjective outcomes
of possible decisions and integrating that information
with memory, perspective frames, and
reward schedules into higher-level strategies.
Perceived value, that differs across individuals, may update
statistical assessment of the environment
to predict delayed reward opportunities in the future.
In doing so, the PMC may continuously adapt to changes
in both the external environment and its internal representation
to enable strategic behavior in volatile environments.

\subsection{The prefrontal cortex: stimulus-value association and action consideration}
Analogous to the PMC,
the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
supra-sensory processes
across time, space, and content domains to
exert top-down control.
In comparison to the PMC, however,
the dmPFC may be closer to a
``mental sketchpad'' \citep{goldman1996prefrontal}, as it
potentially subserves the de-novo generation and binding
of meaning representations instructed by stored semantics and memories.

Generally,
patients with neurological lesions in the prefrontal cortex
are known to struggle with
adaption to novel situations and stimuli
\citep{stuss1986frontal}.
The dmPFC may subserve inference, representation, and assessment
of one's own and other individuals' action considerations.
Specifically, neural activity in the human dmPFC
reflected expectations about other peoples' actions and
errors in these predictions.
Neural activity in the dmPFC indeed explained the proficiency decline
of inferring other peoples' thoughts in the elderly \citep{moran2012social}.
Some dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
In fact, the topographically neighboring dorsal anterior cingulate cortex
has also been linked to computing values and efforts of
persisting a behavioral plan versus switching the
environmental context in several lesion studies \citep{kolling2016value}.
%
Such highly abstract neural computations necessarily rely on the
generation of probabilistic internal information drawing from
episodic memory retrieval, generative construction processes,
and explicit knowledge of the external world.
%
According to computational neuroimaging experiments,
dmPFC activity preferentially models stimulus-value associations of
possible actions that are not actually executed,
whereas vmPFC activity models value in the environment that leads performed actions
\citep{nicolle2012agent}.

Comparing to the dorsomedial prefrontal DMN,
the ventromedial portion of the prefrontal cortex (vmPFC) probably subserves
subjective-value-related evaluative processes and
risk estimation of self-relevant environmental stimuli.
This DMN node is more closely associated with
orchestrating adaptive behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on sophisticated value representations
\citep{doherty2015structure}.
Quantitative lesion findings across 344 human individuals confirmed
a substantial impairment in value-based decision making
\citep{glascher2012lesion}.
Indeed,
the vmPFC is preferentially connected with reward-related and limbic areas.
The vmPFC is well known to have monosynaptical connections
with the NAc
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Congruently, the gray-matter volume of the vmPFC and NAc
correlated with indices of value-guided behavior and reward attitudes
in humans
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals from dopaminergic pathways
\citep{schultz1998predictive}
that not only modulate motivated behavior towards basic survival needs but also
subserve RL in humans more broadly \citep{doherty2015structure}.
This is consistent with diffusion MRI tractography in humans and monkeys
\citep{croxson2005quantitative} that
quantified the NAc to
be substantially more likely connected to the vmPFC than dmPFC in both species.
Two functional connectivity analyses in humans strongly connected
the vmPFC with the NAc, hippocampus (HC),
and PMC.
%
The vmPFC is often proposed to be involved in (external) emotional
reactions and own (visceral) arousal.
Real or imagined bodily states could be mapped in the vmPFC
as a bioregulatory disposition governing cognition
and decision making \citep{damasio1996somatic}
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
valuation
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in the vmPFC.
Such a ``cognitive map'' of the action space was argued to encode
the current task state even when states are unobservable from sensory input,
which was shown to be critical for behavior \citep{Schuck2016}.
% Additionally,
% independent whole-brain analyses from structural
% neuroimaging studies related the gray-matter volume of the vmPFC,
% more consistently than any other
% brain region, to indices of
% social competence and social network complexity, which are probably
% among the most complicated decision that humans and monkeys take
% \citep{behrens2009computation}.

@@

While the HC
is traditionally believed to allow remembering the past,
there is now increasing evidence for a role
in constructing mental models in general
\citep{maguire2016, schacter2007remembering, gelbard2008internally, Javadi2017}.
Its recursive anatomical architecture
may be specifically designed to allow reconstructing
entire episodes of experience from memory fragments.
Indeed,
hippocampal damage is
not only associated with an impairment in reexperiencing the past (i.e., amnesia),
but also imagining about one's own future and
 new fictitious experiences more broadly \citep{hassabis2007patients}.
Mental scenes created by HC-lesioned patients exposed a lack of
spatial integrity, richness in detail, and overall coherence.
%
Single-cell recordings in the animal HC revealed
some constantly active neuronal ensembles whose firing coincided with
distinct locations in space during environmental navigation.
London taxi drivers, individuals with high performance in spatial navigation,
were shown to exhibit increased gray-matter volume in the
HC \citep{maguire2000navigation}.
Indeed, when an animal is choosing between alternative
paths, the corresponding neuronal populations in the HC
spike one after another  \citep{johnson2007neural}.
These neuronal patterns appear to also indicate upcoming behavior
\citep{pfeiffer2013hippocampal}.

But encoding and generative reconstruction in the HC extends
beyond mere spatial knowledge of the environment.
Based on large-scale recordings of hippocampal neuronal populations,
complex spiking patterns can be followed across extended periods including
their modification of input-free self-generated patterns
after environmental events \citep{buzsaki2004large}.
Specific spiking sequences, that were elicited in experimental task conditions,
have been shown to be reenacted spontaneously during
quiet wakefulness and sleep \citep{hartley2014space, o2010play}.
Moreover, spike sequences measured in hippocampal place cells of rats
featured re-occurrence directly after experimental trials
as well as directly before upcoming experimental trials \citep{diba2007forward}.
Such hippocampal ensembles burt during rest and sleep
have been proposed to be critical in communicating local information
to the neocortex for long-term storage, potentially also in the nodes of the DMN.
These HC-subserved mechanisms
probably contribute to
recollecting autobiographical memory episodes, as well as
reexperiencing or newly generating mental scenarios
\citep{hassabis2007patients}.

The HC thus orchestrates experience of environmental aspects for
consolidations based on reenactment and for integration into
rich mental scene construction \citep{deuker2016event, bird2010establishing}.
In this way, the HC may influence
ongoing perception in the current environment
\citep{maguire2016}.
% (TODO: relate to self-play in deepmind Go).
% (TODO: Kai had an interesting idea: the deja-vu effect
% could be when you already accidentally perturbed past
% such that it really occures in reality later.)
